### 2019-11-12 manuscript/manuscript.rmd
update results in manuscript

---
csl: plos.csl
output:
  github_document: default
  pdf_document: default
  html_document: default
bibliography: manuscript.bib
---

@@

Yuanning Li^1^, Xingxing Shen, Benjamin Evans^2^, Antonis Rokas, and Casey W. Dunn^1^*

@@

\* Corresponding author, casey.dunn@yale.edu

@@

Phylogenomic analyses with genome-scale data have the potential to revolutionize our understanding of the tree of life. However, over the past decade there has been considerable debate about the position of the root of the animal phylogeny, with Ctenophora-sister (comb jelly) and Porifera-sister (sponges) (Fig XXOverview) emerging as the two primary hypotheses. Historically, there was little debate about the root of the animal tree of life and Porifera-sister was widely accepted though rarely tested since . In contrast to the lack of debate about the position of Porifera, there has long been uncertainty about the relationship of Ctenophora to other animals [@Wallberg:2004ws]. An understanding of the first diverging animal lineage is critical for scientists to reconstruct events that occurred  early in animal evolution, especially for complex structures such as cell types, nervous systems, and developmental patterns. However, a wide consensus of animal root position has not been reached despite multiple efforts.

@@

![](figures/Figure_overview.png)
**Fig XXOverview.** (A) The Ctenophora-sister hypothesis posits that there is a clade (designated by the orange node) that includes all animals except Ctenophora, and that Ctenophora is sister to this clade. (B) The Porifera-sister hypothesis posits that there is a clade (designated by the green node) that includes all animals except Porifera, and that Porifera is sister to this clade. Testing these hypotheses requires evaluating the support for each of these alternative nodes. (C) A ctenophore. (D) A sponge. (E)

@@

While the exchangeability matrix describes the relative rate of different changes between amino acids, the actual rate can be further scaled. There are couple approaches that have been used in the studies considered here:

@@

- Empirical site homogeneous. The frequency of each amino acid is observed from the matrix under consideration and applied to homogeneously to all sites in the matrix.

- Estimated site homogeneous. The frequency of each amino acid is inferred along with other model parameters, under the assumption that it is the same at all sites.

@@

Models can be assembled by selecting different options for all these different components. The models that are applied in practice area heavily influenced by engineering and computational costs, as well as convention. For example, on the questions considered here F81 and GTR exchangeability matrices have only been used in combination with CAT site heterogeneous models of equilibrium frequency. LG and WAG exchangeability matrices have only been used with site homogeneous estimates of equilibrium frequency. This is further confused by the abbreviations that are used for models. Papers often discuss CAT and WAG models as if they are exclusive, but these particular terms apply to non-exclusive model components-- CAT refers to variation across sites and WAG a particular exchangeability matrix. CAT is generally shorthand for F81+CAT and WAG is shorthand for WAG+homogeneous equilibrium frequency estimation. One could, though, run a WAG+CAT model.

To avoid confusion on this point, we always specify the exchangeability matrix first, followed by modifiers that describe accommodation of heterogeneity in equilibrium frequencies (*e.g.*, CAT) or rate (*e.g.*, Gamma). If there are no modifiers, then it is implied that site homogeneous models are used.

@@

Whole-genome and transcriptomic sequencing is leading to exciting new data matrices for phylogenomic analyses, potentially including hundreds or thousands of genes. Although the number of sampled genes might be different across studies, it is reasonable to expect that the sampled genes should be conserved across animal genomes and most data matrices should share a core set of orthologs. Surprisingly, gene correspondance across different data matrices has been rarely assessed since most studies relied on their own bioinformatic pipeline for determination of orthology genes (OGs).

@@

Outgroup selection (the taxa used to root the ingroup taxa) can potentially affect phylogenetic inference of the root position of animal phylogeny. Several studies have progressively removed more distantly related outgroup taxa to test the effect of outgroup selection to ingroup topology . Three progressively more inclusive clades have often been investigated, including Choanomalia (Metazoa plus most closely related Choanoflagellata), Holoza (Choanimalia plus more distantly related holozoans) and Opisthokonta (Holozoa + Fungi). It has been suggested the inclusion of more distant related (e.g. Opisthokonta) outgroup taxa tend to recover Ctenophora-sister using both site-homogeneous and site-heterogeneous models, whereas high support of Porifera-sister often recovered when site-heterogeneous models were used. In contrast of outgroup selection, sensitivity to ingroup sampling has received less attention than sensitivity to outgroup sampling. This may be because results have tended to be more sensitive to outgroup sampling.

**Fig XXOutgroup.** The animals and their outgroups, showing the three progressively more inclusive clades Choanimalia, Holozoa, and Opisthokonta.

@@

Compositional heterogeneity (lineage-specific differences in amino acid frequencies) was long thought has a major effect in resolving the animal tree of life, especially for the position of animal root. Feuda et al. 2017 has proposed to recode the full set of twenty amino acids into six groups of amino aicds that have more frequent evolutionary changes within groups than between groups based on different exchangeability matrices. By discarding information on which specific amino acid is found at each site in each species and instead focusing on which groups those amino acids belong to, they sought to reduce the impact of differences between species in which particular amino acids within those groups are most frequent. By reanalyzing Chang et al. 2015 and Whelan et al. 2015, they found that Porifera sister was favored under all recoding strategies using GTR-CAT models. However, the performance of the recoding methods has never been tested empirically against non-recoding method in the root of the animal phylogeny, and a recent simulation study has suggested that non-recoding approaches significantly outperformed 6-state recoding strategies. Thus, it is important to evaluate the performance of data recoding method verses non-recoding approaches in empritical animal-root data matrices.

## Overview of published analyses

@@

We synthesize data and results from all previous phylogenomic analyses that tested Ctenophora-sister and Porifera-sister, including 17 data matrices from 12 studies. For taxon composition, several data matrices have relatively skewed taxon sampling, with more than 50% taxa sampled in bilaterians compared to other major animal lineages. These skewed data matrices includes two studies constructed from EST dataset and one study from the sequenced whole animal genome. The rest data matrices have relative balanced taxon composition, although the the number of sampled taxon varies a lot. Since increase taxon sampling is broadly accepted to improve phylogenetic inference an aid in the placement of animal root position, it should be noted that Whelan2017 and Simion2017 have much broader taxon sampling than other data matrices, with Simion2017 has the most taxon sampling and Whelan2017 contains greater ctenophore taxon sampling.

@@

**Fig XXTaxon_composition.** Each of the primary matrices considered here, color coded by taxon sampling. Horizontal size is proportional to the number of genes (XXOr should it be sites?) sampled, vertical size to the number of taxa sampled.

### Matrix gene composition

Gene sampling plays an essential role in phylogenomic analyses, and most often rely on the concatenation of sampled genes to tackle the incongruence of the topology among individual gene trees. The basic idea is that the concatenation assumes that phylogenetic signal from genes that do not share the species phylogeny will be overwhelmed by the signal from the majority of genes whose genealogy mirrors that of the species evolutionary history [@lanier2012recombination]. However, large phylogenomic data matrices potentially contain genes with conflicting or weak phylogenetic signal that can lead to incongruence phylogenomic results, especially when analyzing deeper node represents ancient divergence in the tree [@salichos2013inferring]. Thus, it is important to evaluate the gene composition and matrix overlap across difference data matrices to explore.

We first used the Metazoa BUSCO dataset to evaluate gene composition from all data matrices and the result is unexpected. The animal BUSCO dataset contains 978 single-copy OGs (identified from 64 animal taxa) that are most likely present in the common ancestor of animal genome, and are commonly used to assess the completeness of animal genome or transcriptomic assembly and frequently used as phylogenetic markers. Thus, it is reasonable to expect that these conserved genes should present at a higher percentage than non-BUSCO genes in most data matrices given their nature of conserveness and high-expression in transcriptomes. However, to our surprise, we found that OGs used in all data matrices comprised less than 50%, even with several data matrices contain less than 20 percent of BUSCO genes.

We furthur

@@

p2 = ggplot( matrix_summary ) + geom_point( aes(x=matrix, y=n_busco_partitions/n_partitions*100, col=manuscript) ) + 	theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))
count =  busco_results %>% select( matrix,  Busco_Id ) %>% distinct(  )
reference_busco=read_tsv("/animal_root/manuscript/busco.tsv")

Simion2017_busco <- count[grep("Simion2017", count$matrix), ] %>% select(Busco_Id)
names(Simion2017_busco)[1]<-"Simion2017"

@@

### Matrix overlap

Previous reanalyses have mainly focused on ourgroup selection and/or model fits in different data matrices, and the matrix overlap across different studies has never been investigated. Our initial intent of comparing matrix overlap across matrices was to constructed new matrices with shared taxon and gene sampling to test specific hypotheses about differences in support. The hypothesis is that although the bioinformatics pipeline used for OG determination is differt, most larger data matrices should share a core set of OGs. However, our results show that surprisingly little overlap were identified across all the data matrices from different studies (Fig XXAlignment overlap).

@@

Here we summarized all the previous phylogenetic analyses from all studies and the results are shown in Fig. support_published_analyses and Table perviouly_published_analyses.tsv. We found that all phylogenomic analyses strongly support of Ctenophora-sister once site-homogeneous model were used, no matter what outgroup choice, phylogenetic inference, recoding method, data partitioning are used. We found that support for Porifera-sister increases with the exclusion of more distantly related outgroups and the use of site-heterogeneious CAT models, although many of analyses are not converged or posterior probability is relatively low using PhyloBayes. It should be noted that most support of Porifera-sister are from analyses using Poisson + CAT and recoding methods with GTR + CAT models.

@@

One of the challenges of interpreting support for the placement of the animal root across studies is that different programs, software versions, and settings have been used across studies, and phylogenetic analysis decisions have been approached in very different ways. Here we reanalyze the primary matrices from each study under consistent conditions with IQtree under a panel of evolutionary models. We selected this tool because it has greater model flexibility, computational time and accuracy than other tools in maximum likelihood framework.

We first tested a variety of models for each matrix, and inferred support under the best-fit model from IQtree. We then analyzed every matrix under a panel of standard models, including WAG+G, GTR+G, Poisson+G.

@@

Site heterogeneity in equilibrium frequency has been a major concern in tests of Ctenophora-sister and Porifera-sister. This has been addressed with CAT models. iqtree provides a new family of C models that also address site heterogeneity. Given the extensive computational cost and concerns about overparameterization of CAT models, we compared iqtree C results to CAT results for a subset of matrices to see if they give consistent results. This would be of technical interest because it would reduce the cost of accommodating compositional heterogeneity in future analyses.

@@

All files associated with this analysis are available at https://github.com/caseywdunn/animal_root

@@

Whelan *et al.* is the only study to evaluate performance of site-heterogeneous models and site-homogeneous model with data partitioning under the simulation framework. The simulation results suggested that Poisson+CAT model consistently performed worse than other models in simulation datasets. More importantly, the authors also showed that both Poisson + CAT and GTR +  CAT models could overestimated substitutional heterogeneity in many cases. They also reanalyzed datasets from Philippe 2009 and Nosenko 2013 using both CAT models and data partitioning with site-homogeneious model. The results indicated that Poisson + CAT model tends to recover less accurate trees and both GTR + CAT and data partitioning strongly supported Ctenophora-sister in reanalyses. The authors also concluded:

>>>>

---
bibliography: manuscript.bib
csl: plos.csl
output:
  word_document: default
  html_document: default
  pdf_document: default
  github_document: default
---

@@

Yuanning Li^1,3^, Xingxing Shen^3^, Benjamin Evans^2^, Antonis Rokas^3^, and Casey W. Dunn^1^*

@@

^3^Department of Biological Sceince, Vanderbilt University
\* Corresponding author, casey.dunn@yale.edu

@@

Phylogenomic analyses with genome-scale data have the potential to revolutionize our understanding of the tree of life. However, over the past decade, there has been considerable debate about the position of the root of the animal phylogeny, with Ctenophora-sister (comb jelly) and Porifera-sister (sponges) (Fig XXOverview) emerging as the two primary hypotheses. Historically, there was little debate about the root of the animal tree of life and Porifera-sister was widely accepted though rarely tested. In contrast to the lack of debate about the position of Porifera, there has long been uncertainty about the relationship of Ctenophora to other animals [@Wallberg:2004ws]. An understanding of the first diverging animal lineage is critical for scientists to reconstruct events that occurred  early in animal evolution, especially for complex structures such as cell types, nervous systems, and developmental patterns. However, a wide consensus of the position of animal has not been reached despite multiple efforts.

@@

![](figures/Figure_overview.png)
**Fig XXOverview.** (A) The Ctenophora-sister hypothesis posits that there is a clade (designated by the orange node) that includes all animals except Ctenophora, and that Ctenophora is sister to this clade. (B) The Porifera-sister hypothesis posits that there is a clade (designated by the green node) that includes all animals except Porifera, and that Porifera is sister to this clade. Testing these hypotheses requires evaluating the support for each of these alternative nodes. (C) The animals and their outgroup choice, showing the three progressively more inclusive clades Choanimalia, Holozoa, and Opisthokonta (designated by the blue node). (D) A ctenophore. (E) A sponge.

@@

While the exchangeability matrix describes the relative rate of different changes between amino acids, the actual rate can be further scaled. There are a couple approaches that have been used in the studies considered here:

@@

- Empirical site-homogeneous. The frequency of each amino acid is observed from the matrix under consideration and applied to homogeneously to all sites in the matrix.

- Estimated site-homogeneous. The frequency of each amino acid is inferred along with other model parameters, under the assumption that it is the same at all sites.

@@

Models can be assembled by selecting different options for all these different components. The models that are applied in practice area heavily influenced by engineering and computational costs, as well as convention. For example, on the questions considered here, F81 and GTR exchangeability matrices have only been used in combination with CAT site heterogeneous models of equilibrium frequency. LG and WAG exchangeability matrices have only been used with site homogeneous estimates of equilibrium frequency. This is further confused by the abbreviations that are used for models. Papers often discuss CAT and WAG models as if they are exclusive, but these particular terms apply to non-exclusive model components-- CAT refers to variation across sites and WAG a particular exchangeability matrix. CAT is generally shorthand for F81+CAT and WAG is shorthand for WAG+homogeneous equilibrium frequency estimation. One could, though, run a WAG+CAT model.

To avoid confusion on this point, we always specify the exchangeability matrix first, followed by modifiers that describe the accommodation of heterogeneity in equilibrium frequencies (*e.g.*, CAT) or rate (*e.g.*, Gamma). If there are no modifiers, then it is implied that site homogeneous models are used.

@@

High-throughput sequencing is leading to exciting new data matrices for phylogenomic analyses, potentially including hundreds or thousands of genes (*e.g.*, [@whelan2017ctenophore; @simion2017large]). Although the number of sampled genes might be different across studies, it is reasonable to expect that the sampled genes should be conserved across animal genomes and most data matrices should share a core set of orthology genes (OGs) (*e.g.*, BUSCO genes). Surprisingly, gene composition and matrice overlap across different studies has been rarely assessed, since most studies relied on their own bioinformatic pipeline for determination of OGs.

@@

Outgroup selection (the taxa used to root the ingroup taxa) can potentially affect the phylogenetic inference of the root position of animal phylogeny (**Fig XXOverview.C**). Several studies have progressively removed more distantly related outgroup taxa to test the effect of outgroup selection to ingroup topology [@ryan2013genome; @pisani2015genomic]. Three progressively more inclusive clades have often been investigated, including Choanomalia (Metazoa plus most closely related Choanoflagellata), Holoza (Choanimalia plus more distantly related holozoans) and Opisthokonta (Holozoa + Fungi). It has been suggested the inclusion of more distant related (e.g. Opisthokonta) outgroup taxa tend to recover Ctenophora-sister, whereas high support of Porifera-sister often recovered only when site-heterogeneous CAT models were used. In contrast of outgroup selection, sensitivity to ingroup sampling has received less attention than sensitivity to outgroup sampling. This may be because results have tended to be more sensitive to outgroup sampling.

@@

Compositional heterogeneity (lineage-specific differences in amino acid frequencies) was long thought has a major effect in resolving the animal tree of life, especially for the animal root position. Feuda et al. 2017 [@feuda2017improved] has proposed to recode the full set of twenty amino acids into six groups that have more frequent evolutionary changes within groups than between groups based on different exchangeability matrices. By discarding information on which specific amino acid is found at each site in each species and instead focusing on which groups those amino acids belong to, they sought to reduce the impact of differences between species in which particular amino acids within those groups are most frequent. By reanalyzing Chang et al. 2015 [@chang2015genomic] and Whelan et al. 2015 [@whelan2015error], they found that Porifera-sister was strongly favored under all recoding strategies using GTR-CAT models [@feuda2017improved]. However, the performance of the recoding methods has never been tested empirically against the non-recoding method, and a recent simulation study has suggested that non-recoding approaches significantly outperformed 6-state recoding strategies [@hernandez2019six]. Thus, it is important to evaluate the performance of data recoding method verses non-recoding approaches in past empritical matrices .

## Overview of published analyses
Here, we synthesize data and results from all previous phylogenomic analyses that tested Ctenophora-sister and Porifera-sister, including 17 data matrices from 12 studies (Table 1).

Table 1.
```{r overview data matrices}
	kable(datasets)

@@

Taxa sampling plays a critical role in phylogenetic inference, especially for the phylogenetic position that needs to be explored. However, several data matrices have relatively skewed taxon sampling, with more than 50% taxa sampled in bilaterians compared to other major animal lineages (Fig XXTaxon_composition). These skewed data matrices include two studies constructed from ESTs (Dunn2008 [@Dunn:2008ky] and Henjol2009 [@hejnol2009assessing]) and one study from whole-genome data (Borrowiec2015 [@borowiec2015extracting] ). Importantly, some earlier studies only include less than five ctenophore species. The rest data matrices have relatively balanced taxon composition, although the total number of sampled taxon varies a lot. In general, increasing taxon sampling is broadly accepted to improve phylogenetic inference and aid in the placement of animal root position, and it should be noted that Whelan2017 [@whelan2017ctenophore] and Simion2017 [@simion2017large] have much broader taxon sampling than other data matrices: Simion2017 has the most taxon sampling, and Whelan2017 contains greater ctenophore taxon sampling.

@@

**Fig XXTaxon_composition.** Each of the primary matrices considered here, color coded by taxon sampling. Horizontal size is proportional to the number of genes sampled, vertical size to the number of taxa sampled.

### Matrix gene composition and overlap

Gene sampling plays an essential role in phylogenomic analyses, and most often rely on the concatenation of sampled genes to tackle the incongruence of the topology among individual gene trees. The basic idea is that the concatenation assumes that phylogenetic signal from genes that do not share the species phylogeny will be overwhelmed by the signal from the majority of genes whose genealogy mirrors that of the species evolutionary history [@lanier2012recombination]. However, large phylogenomic data matrices potentially contain genes with conflicting or weak phylogenetic signal that can lead to incongruence phylogenomic results, especially when analyzing deeper node represents ancient divergence in the tree [@salichos2013inferring]. Thus, it is important to evaluate the gene composition and matrix overlap to explore congruence of gene sampling across different matrices. It should be noted that early studies relied on EST via Sanger sequencing contains large missing data compared to other ones.

We first used the Metazoa BUSCO dataset to evaluate gene composition from all data matrices and the result is unexpected. The animal BUSCO dataset contains 978 single-copy OGs (identified from 64 animal taxa) that are most likely shared by most animal genomes, and are commonly used to assess the completeness of genome or transcriptomic assemblies. Morever, the BUSCO dataset also are frequently used as phylogenomic markers in a varietly of studies. Thus, it is reasonable to expect that these conserved genes should present at a higher percentage than non-BUSCO genes in most matrices given their nature of conserveness and high-expression properties. However, to our surprise, we found that gene partitions from all matrices comprised less than 50% of BUSCO genes, even with several data matrices contain less than 20 percent (Fig. XXBUSCO_annotations).

We furthered BLAST gene partitions to SwissProt database

Previous reanalyses have mainly focused on ourgroup selection and/or model fits in different data matrices, and the matrix overlap across different studies has been rarely investigated. Our initial intent of comparing matrix overlap across matrices was to construct a new matrice with shared taxon and gene sampling to test specific hypotheses about differences in support. The hypothesis is that although the bioinformatics pipeline used for OG determination is different, most larger data matrices should share a core set of OGs. However, our results show that surprisingly little overlap across all the data matrices from different studies (Fig XXAlignment overlap). These results indicated that different studies essentially sampled different regions of the genomes with little overlap. One thing should be noted that

@@

p2 = ggplot( matrix_summary ) + geom_point( aes(x=matrix, y=n_busco_partitions/n_partitions*100, col=manuscript) ) + 	theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))

@@

@@

Here we summarized all the 167 previous phylogenetic analyses from all studies (Fig. support_published_analyses and Table perviouly_published_analyses). We found that all past phylogenomic analyses strongly support of Ctenophora-sister once site-homogeneous models were used, no matter what outgroup choice, phylogenetic inference, recoding methods are used. We found that support for Porifera-sister increases with the exclusion of more distantly related outgroups with the use of site-heterogeneious CAT models. It should be noted that the convergence were not reached or posterior probability is not well supported in many BI analyses. Moreover, the strong support of Porifera-sister are mainly from analyses using Poisson+CAT and recoding methods with site-heterogeneous models. Moreover, only a few past analyses with the most complex GTR+CAT model are converged. However, for the ones that are converged with GTR+CAT model, Porifera-sister is not always supported since several analyses strongly in favor Ctenophora-sister when choanozoa are used as outgroups.

The use of site-heterogeneous CAT is an infinite mixture model has long been suggested as the best approach to account for systematic errors, particular for data saturation, especially in the case of long-branch attraction (LBA) (Lartillot and Philippe, 2004). The GTR-CAT model often received the best-model fit from cross validation or posterior predictive analyses compared to site-homogeneous or Poisson-CAT models. However, the GTR+CAT model is extremely computational demanding and not feasible in large matrices so many studies only include Poisson+CAT models in their analyses. In contrast, a recent study has suggested that Poisson+CAT model did not outperform site-homogeneous with data partitioning or GTR + CAT models and in some cases can infer incorrect topology based on the simulation analyses. The author also raised the issue of the potential overparameterization in CAT models in some cases. Nevertheless，the debate over the site-heterogeneous vs. site-homogeneous models and/or inclusion or exclusion more distantly related outgroups, constitutes the probably one of the most controversial nodes in the animal tree of life.

@@

One of the challenges of interpreting support for the placement of the animal root across studies is that different programs, software versions, and settings have been used across studies, and phylogenetic analysis decisions have been approached in very different ways. Here we reanalyze the primary matrices from each study under consistent conditions with IQtree under a panel of evolutionary models. We selected this tool because it has greater model flexibility, computational time and accuracy than other tools in maximum likelihood (ML) framework. Importantly, we also include a CAT-like, site-heterogeneous C10 - C60 (C) models that are implemented in both ML framework in our analyses and it is interesting to compare results between C model to CAT models since it has never been used in any matrices analyzed here.

We first tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C models using ModelFinder in IQtree. We found in all cases, C60+ WAG/LG models fit these matrices better than the site-homogeneous models under BIC criteria and we then inferred support under the best-fit model from IQtree. It should be noted that we didn't include GTR + C models in model testing since it is computational too expensive. The GTR+C models requires to estimate up to over 10,000 (e.g. GTR+C60 contains 234*60 parameters) free parameters inferred from matrices, whereas only several hundred parameters needs to estimate in site-homogeneous + C models. We then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG+G, GTR+G and Poisson+C60+F+G. Moreover, we also used the best-fit model identified above with the removal of all distantly related outgroups and only include Choanoflagellata as outgroups.

With the exception of Moroz2014 matrix that supports neither Porifera or Ctenophora-sister, every analysis conducted herein choice strongly supported the ctenophore-sister hypothesis, even with site-heterogeneous C models with the most closely related outgroups.

@@

Site heterogeneity in equilibrium frequency has been a major concern in tests of Ctenophora-sister and Porifera-sister. This has been addressed with CAT models. iqtree provides a new family of C models that also address site heterogeneity. Given the extensive computational cost and concerns about overparameterization of CAT models, we compared iqtree C results to CAT results for a subset of matrices to see if they give consistent results. This would be of technical interest because it would reduce the cost of accommodating compositional heterogeneity in future analyses. For computational efficiency, we first only applied Poisson+CAT models in every matrice with only Choanoflagellata as outgroup. It should be noted that most Phylobayes runs were converged, several large matrices have not reached convergence after at least a month's computational time.

Interestingly, we found strong supports for both hypotheses when using Poisson-CAT model in Choanimalia matrices. Consistent with previous study, we recovered the strong support of Porifera-sister in matrices of Philippe2008, Ryan2013, and Whelan2015. Moreover, we also found a strong support of Porifera-sister in two Whelan2017 matrices and a weaker support of subsampled matrice of Simion2017. Similar to ML analyses, we found no strong support for either hypothesis in Moroz2013_3d matrice. For the matrices have conflicted results between Poisson-CAT and C models, we further ran Poisson+CAT models with inclusion of more distantly related outgroups to evaluate how outgroup choice will affect phylogenetic inference under CAT model. Consistently with past study, we found strong support of Ctenophora-sister in all anaylzed Opisthokonta matrices, indicating that Ctenophora-sister is also favored in all matrices with more distantly related taxa are used as outgroups.

###

@@

All files and bioinformatic commmands associated with this analysis are available at https://github.com/caseywdunn/animal_root

@@

Whelan *et al.* is the only study to evaluate performance of site-heterogeneous models and site-homogeneous model with data partitioning under the simulation framework. The simulation results suggested that Poisson+CAT model consistently performed worse than other models in simulation datasets. More importantly, the authors also showed that both Poisson + CAT and GTR +  CAT models could overestimated substitutional heterogeneity in many cases. They also reanalyzed datasets from Philippe 2009 and Nosenko 2013 using both CAT models and data partitioning with site -homogeneious model. The results indicated that Poisson + CAT model tends to recover less accurate trees and both GTR + CAT and data partitioning strongly supported Ctenophora-sister in reanalyses. The authors also concluded:
